### 2023-02-05 drafting_supplementary_analysis.Rmd

---
title: "Supplementary Analyses"
author: "Mariola Paruzel-Czachura, Cillian McHugh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: "resources/bib/My Library.bib"
csl: "resources/bib/apa.csl"
output:
  bookdown::html_document2:
    fig_caption: yes
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 4
    number_sections: false
editor_options: 
  chunk_output_type: console
---

<style>
  .superbigimage{
      overflow-x:scroll;
      white-space: nowrap;
  }

  .superbigimage img{
     max-width: none;
  }

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = FALSE)
```

```{r}
rm(list = ls())
library(tidyverse)
library(plyr)
library(psych)
library(skimr)
library(corrplot)
library(Hmisc)
#library(psy)
library(afex)
#devtools::install_github("cillianmiltown/R_desnum")
library(desnum)
library(papaja)
library(purrr)
#install.packages("interactions")
library(interactions)
library(kableExtra)
#devtools::install_github("crsh/papaja")
library("papaja")
library(lme4)
library(car)
library(ggplot2)
library(nlme)
#install.packages("reshape")
library(reshape)
library(corx)
library(lsr)
#install.packages("Cairo")
#library(Cairo)
library(foreign)
library(ltm)
library("ggpubr")
#install.packages("rworldmap")
library(rworldmap)
#devtools::install_github("rensa/ggflags")
library(ggflags)
#devtools::install_github("schliebs/ggoxford")
library(ggoxford)
library(ggforce)
library(ggbeeswarm)
library(countrycode)
```

```{r}

#source("set_up_data_calculate_means.R")
load("data/data_with_means.RData")

df$sex3 <- as.factor(df$sex1)
df$sex3 <- dplyr::recode(df$sex3,
               "1"="male"
              ,"2"="female"
              ,"3"="other")

df$country_lower <- tolower(df$country)
df$country_name <- countrycode(df$ISO3, origin = "iso3c", destination = "country.name")
```

# H1 Gender Differences

## Internalization

```{r}

ps <- function(y){
  y <- round(y, digits = 3)
  if(as.numeric(sqrt( y*y) ) <.001) print(paste0("<", " ", ".001","**"), quote = FALSE)
  else if(as.numeric(sqrt( y*y) ) <.05) print(paste0(sub("^(-?)0.", "\\1.", sprintf("%.3f", 
        y, quote = FALSE)),"*"))
  else print(sub("^(-?)0.", "\\1.", sprintf("%.3f", 
        y)))}

#### ensure correct data is loaded ####
x <- df[which(df$sex3=="male"|df$sex3=="female"),]
x$sex3 <- droplevels(x$sex3)

#### remove any missing data #### 
sum(is.na(x$moral_identity_internal))
x <- x[which(is.na(x$moral_identity_internal)==FALSE),]
table(x$country_name,x$sex3)
x$country_name

#### group by country ####
x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )

#### identify countries with small samples
male_length <- function(x){sum(x$sex3=='male')}
males <- lapply(x, male_length)
female_length <- function(x){sum(x$sex3=='female')}
females <- lapply(x, female_length)
names(females)

#### generate a vector containing all countries with less than 10 observations for either males or females ####
exclusions <- c(names(males[males<2]),
  names(females[females<2]))

#### re-load ungrouped data ####
x <- df[which(df$sex3=="male"|df$sex3=="female"),]
x$sex3 <- droplevels(x$sex3)

# and remove missing again
sum(is.na(x$moral_identity_internal))
x <- x[which(is.na(x$moral_identity_internal)==FALSE),]

#### Filter out any countries to be excluded ####
x <- subset(x, !(country_name %in% exclusions))

#### group by country for analysis ####
x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )

ttest_fun <- function(x){
  t.test(moral_identity_internal ~ sex3, data = x)
}
all_t_tests <- lapply(x, ttest_fun)

# Means and Standard Deviations
means_fun <- function(x){
  mean(x$moral_identity_internal)
}
lapply(x, means_fun)
sd_fun <- function(x){
  sd(x$moral_identity_internal)
}
lapply(x, sd_fun)

N_fun <- function(x){
  length(x$sex3)
}
total_Ns <- lapply(x, N_fun)

# Statistics for table
# Cohen's d
cohens_d_fun <- function(x){
  cohen.d(x$moral_identity_symbolic, x$sex3)
}
all_ds <- lapply(x, cohens_d_fun)
show_d_fun <- function(x){
  x$cohen.d[2]
}
ds_for_table <- lapply(all_ds, show_d_fun)

# t statistic
all_t_stats <- function(t){
  t$statistic
}
ts_for_table <- lapply(all_t_tests, all_t_stats)

# df
all_t_df <- function(t){
  t$parameter
}
df_for_table <- lapply(all_t_tests, all_t_df)

# p value
all_t_ps <- function(t){
  t$p.value
}
ps_for_table <- lapply((lapply(all_t_tests, all_t_ps)), ps)

sum(unlist(lapply(all_t_tests, all_t_ps))<.05)

#### means for Males ####
# re-load ungrouped data #
x <- df[which(df$sex3=="male"
              #|df$sex3=="female"
              ),]
x$sex3 <- droplevels(x$sex3)

# and remove missing again
sum(is.na(x$moral_identity_internal))
x <- x[which(is.na(x$moral_identity_internal)==FALSE),]
x <- subset(x, !(country_name %in% exclusions))

x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )
male_means <- lapply(x, means_fun)
male_sds <- lapply(x, sd_fun)
male_Ns <- lapply(x, N_fun)

#### means for Females
# re-load ungrouped data #
x <- df[which(df$sex3=="female"
              #|df$sex3=="female"
              ),]
x$sex3 <- droplevels(x$sex3)

# and remove missing again
sum(is.na(x$moral_identity_internal))
x <- x[which(is.na(x$moral_identity_internal)==FALSE),]
x <- subset(x, !(country_name %in% exclusions))

x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )
female_means <- lapply(x, means_fun)
female_sds <- lapply(x, sd_fun)
female_Ns <- lapply(x, N_fun)

```

### Internalization Ns

```{r}
#### Total Ns
x <- df[which(df$sex3=="male"
              |df$sex3=="female"
              ),]
x$sex3 <- droplevels(x$sex3)

# and remove missing again
sum(is.na(x$moral_identity_internal))
x <- x[which(is.na(x$moral_identity_internal)==FALSE),]
x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )
total_Ns <- lapply(x, N_fun)

#### Ns for Males 
# re-load ungrouped data #
x <- df[which(df$sex3=="male"
              #|df$sex3=="female"
              ),]
x$sex3 <- droplevels(x$sex3)

# and remove missing again
sum(is.na(x$moral_identity_internal))
x <- x[which(is.na(x$moral_identity_internal)==FALSE),]
x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )
male_Ns <- lapply(x, N_fun)

#### Ns for Females
# re-load ungrouped data #
x <- df[which(df$sex3=="female"
              #|df$sex3=="female"
              ),]
x$sex3 <- droplevels(x$sex3)
# and remove missing again
sum(is.na(x$moral_identity_internal))
x <- x[which(is.na(x$moral_identity_internal)==FALSE),]
x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )
female_Ns <- lapply(x, N_fun)

test <- 
  cbind.data.frame(do.call(rbind, total_Ns)
                 ,do.call(rbind, female_Ns)
                 ,do.call(rbind, male_Ns))

test <- `colnames<-`(test, c("Total N","n Female","N male"))

loop_exclusions_fun <- function(x){
  if(rownames(test)[x] %in% exclusions) print(paste0(rownames(test)[x],"*"),quote = FALSE)
  else print(rownames(test)[x])
  }

rownames(test) <-  unlist(lapply(1:length(rownames(test)), loop_exclusions_fun))
test
test <- test[order(rownames(test) ),]

```

```{r include=TRUE}

apa_table(
  test
  , align = c("c", "c", "c")
  , caption = "Total samples for each country for Internalization analysis"
  , note = "* = sample too small to be included in analysis below")

```

### Internalization Table

```{r}

test <- 
  cbind.data.frame(
     round(do.call(rbind, female_means),digits=2)
    ,round(do.call(rbind, female_sds),digits=2)
    ,round(do.call(rbind, male_means),digits=2)
    ,round(do.call(rbind, female_sds),digits=2)
    ,round(do.call(rbind, ts_for_table),digits=2)
    ,round(do.call(rbind, df_for_table),digits=2)
    ,do.call(rbind, ps_for_table)
    ,round(do.call(rbind, ds_for_table),digits=2)
    )

test <- `colnames<-`(test, c("mean F","SD F","mean M", "SD M", "t","df","p","d"))
test

x <- df[which(df$sex3=="male"
              |df$sex3=="female"
              ),]
x$sex3 <- droplevels(x$sex3)

# and remove missing again
sum(is.na(x$moral_identity_internal))
x <- x[which(is.na(x$moral_identity_internal)==FALSE),]
x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )
total_Ns <- lapply(x, N_fun)

P_R <- x$`Puerto Rico`
P_R <- 
  `rownames<-`(
    `colnames<-`(
      as.data.frame(
        rbind(
          # c("mean F","SD F","mean M", "SD M", "t","df","p","d")
          # ,
          c(
            P_R$moral_identity_internal[which(P_R$sex3=="male")]
            ,"-"
            ,P_R$moral_identity_internal[which(P_R$sex3=="female")]
            ,"-"
            ,"-"
            ,"-"
            ,"-"
            ,"-"))),
      c("mean F","SD F","mean M", "SD M", "t","df","p","d"))
    ,c("Puerto Rico"))

test <- rbind(test[1:49,],P_R,test[50:length(test$`mean F`),])
test <- test[order(rownames(test) ),]

```

```{r include=TRUE}

apa_table(
  test
  , align = c("l", "l", "c", "c", "c", "c", "c", "c", "c", "c")
  , caption = "Gender differences in Internalization for each Country"
  , note = "* = sig. at p < .05; ** = sig. at p < .001")

```

### Internalization Plot

```{r internalizationPlot}

# x <- df[which(df$sex3=="male"|df$sex3=="female"),]
# x$sex3 <- droplevels(x$sex3)
# x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )
# 
# #### identify countries with small samples
# male_length <- function(x){sum(x$sex3=='male')}
# males <- lapply(x, male_length)
# female_length <- function(x){sum(x$sex3=='female')}
# females <- lapply(x, female_length)
# names(females)
# 
# #### generate a vector containing all countries with less than 10 observations for either males or females ####
# exclusions <- c(names(males[males<2]),
#   names(females[females<2]))
# 
# #### re-load ungrouped data ####
# x <- df[which(df$sex3=="male"|df$sex3=="female"),]
# x$sex3 <- droplevels(x$sex3)
# 
# # and remove missing again
# sum(is.na(x$moral_identity_internal))
# x <- x[which(is.na(x$moral_identity_internal)==FALSE),]
# 
# #### Filter out any countries to be excluded ####
# x <- subset(x, !(country_name %in% exclusions))
# 
# df1 <- x
# 
# df1_male <- df1[which(df1$sex3=="male"),]
# df1_female <- df1[which(df1$sex3=="female"),]
# 
# internal_mean_sd <- function(x){
#   z <- x
#   z$mean <- mean(z$moral_identity_internal, na.rm=T)
#   z$sd <- sd(z$moral_identity_internal, na.rm=T)
#   z
# }
# 
# 
# df1 <- rbind.data.frame(
#   internal_mean_sd(df1_female),
#   internal_mean_sd(df1_male)
# )
# 
# df1$sex3 <- dplyr::recode(df1$sex3
#                            , "male" = "Male"
#                            , "female" = "Female")
# 
# g1 <- ggplot(df1,aes(x=sex3,y=moral_identity_internal))+
#   #geom_jitter(height = .1, width = .1, size=.0001, color="gray") + 
#   geom_jitter(shape=16, position=position_jitter(height = .5, width = .45), size=.1, color="lightgrey")+
#   geom_violin(fill = NA) +
#   #ggforce::geom_sina(size=.01)+
#   # geom_sina(shape=16
#   #           , position=position_jitter(height=.25,width=0)
#   #           , size=.000001
#   #           , color="dark grey"
#   #           ,jitter_y=F
#   # )+
#   geom_boxplot(width=0.1, outlier.shape = NA, color='grey')+
#   geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.06,
#                 position=position_dodge(0.05), color="#5a5a5a")+
#   stat_summary(fun=mean, geom="point", shape=16, size=.5)+
#   # stat_summary(geom = "linerange",
#   #   fun.data = mean_sdl, 
#   #   fun.args = list(mult = 1),
#   #   colour = "black"
#   # )+
#   xlab("Gender")+
#   ylab("Internalization")+  
#   facet_wrap(~df1$ISO3, ncol = 9 )+
#   #geom_jitter(shape=16, position=position_jitter(height = .5, width = .45), size=.01, color="dark grey")
#   theme_bw() +
#   theme(panel.border = element_blank(),
#         axis.line = element_line(size = .2),
#         strip.background  = element_blank(),
#         panel.grid = element_blank(),
#         plot.title=element_text(#family="Times",
#           size=12
#         ),
#         #legend.position="right",
#         legend.text=element_text(#family="Times",
#           size=8
#         ),
#         legend.title=element_text(#family="Times",
#           size=10
#         ),
#         axis.text=element_text(#family="Times",
#           colour = "black",
#           size=8
#         ),
#         axis.ticks.x = element_blank(),
#         axis.title=element_text(#family="Times",
#           size=12
#         ),
#         strip.text=element_text(#family = "Times",
#           size = 12
#         ),
#         # strip.background = element_rect(fill = "white"),
#         legend.position="none")
# 
# ggsave("plots/internalization.png", g1, width = 40, height = 25, units = "in", dpi = 300)

```

```{r, InternalizationGender,fig.cap="Gender differences in Internalization for each country",include=TRUE}
#suppressWarnings(print(g1))

```

<style>
#wrapper { width: 100%;; height: 702px; padding: 0; overflow: hidden; border: 1px}
#scaled-frame { width: 250%; height: 900px; border: 1px; }
#scaled-frame {
    zoom: 0.50;
    -moz-transform: scale(0.50);
    -moz-transform-origin: 0 0;
    -o-transform: scale(0.50);
    -o-transform-origin: 0 0;
    -webkit-transform: scale(0.50);
    -webkit-transform-origin: 0 0;
}

@media screen and (-webkit-min-device-pixel-ratio:0) {
 #scaled-frame  { zoom: 1;  }
}
</style>

<div id="wrapper"><iframe id="scaled-frame" src="plots/internalization.png"></iframe></div>

## Symbolization

```{r}

ps <- function(y){
  y <- round(y, digits = 3)
  if(as.numeric(sqrt( y*y) ) <.001) print(paste0("<", " ", ".001","**"), quote = FALSE)
  else if(as.numeric(sqrt( y*y) ) <.05) print(paste0(sub("^(-?)0.", "\\1.", sprintf("%.3f", 
        y, quote = FALSE)),"*"))
  else print(sub("^(-?)0.", "\\1.", sprintf("%.3f", 
        y)))}

#### ensure correct data is loaded ####
x <- df[which(df$sex3=="male"|df$sex3=="female"),]
x$sex3 <- droplevels(x$sex3)

#### remove any missing data #### 
sum(is.na(x$moral_identity_symbolic))
x <- x[which(is.na(x$moral_identity_symbolic)==FALSE),]
table(x$country_name,x$sex3)
x$country_name

#### group by country ####
x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )

#### identify countries with small samples
male_length <- function(x){sum(x$sex3=='male')}
males <- lapply(x, male_length)
female_length <- function(x){sum(x$sex3=='female')}
females <- lapply(x, female_length)
names(females)

#### generate a vector containing all countries with less than 10 observations for either males or females ####
exclusions <- c(names(males[males<2]),
  names(females[females<2]))

#### re-load ungrouped data ####
x <- df[which(df$sex3=="male"|df$sex3=="female"),]
x$sex3 <- droplevels(x$sex3)

# and remove missing again
sum(is.na(x$moral_identity_symbolic))
x <- x[which(is.na(x$moral_identity_symbolic)==FALSE),]

#### Filter out any countries to be excluded ####
x <- subset(x, !(country_name %in% exclusions))

#### group by country for analysis ####
x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )

ttest_fun <- function(x){
  t.test(moral_identity_symbolic ~ sex3, data = x)
}
all_t_tests <- lapply(x, ttest_fun)

means_fun <- function(x){
  mean(x$moral_identity_symbolic)
}
lapply(x, means_fun)
sd_fun <- function(x){
  sd(x$moral_identity_symbolic)
}
lapply(x, sd_fun)

cohens_d_fun <- function(x){
  cohen.d(x$moral_identity_symbolic, x$sex3)
}
all_ds <- lapply(x, cohens_d_fun)
show_d_fun <- function(x){
  x$cohen.d[2]
}
ds_for_table <- lapply(all_ds, show_d_fun)

all_t_stats <- function(t){
  t$statistic
}
ts_for_table <- lapply(all_t_tests, all_t_stats)

all_t_df <- function(t){
  t$parameter
}
df_for_table <- lapply(all_t_tests, all_t_df)

all_t_ps <- function(t){
  t$p.value
}
ps_for_table <- lapply((lapply(all_t_tests, all_t_ps)), ps)

total_sig <- sum(unlist(lapply(all_t_tests, all_t_ps))<.05)

#### means for Males ####
# re-load ungrouped data #
x <- df[which(df$sex3=="male"
              #|df$sex3=="female"
              ),]
x$sex3 <- droplevels(x$sex3)

# and remove missing again
sum(is.na(x$moral_identity_symbolic))
x <- x[which(is.na(x$moral_identity_symbolic)==FALSE),]
x <- subset(x, !(country_name %in% exclusions))

x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )
male_means <- lapply(x, means_fun)
male_sds <- lapply(x, sd_fun)

#### means for Females
# re-load ungrouped data #
x <- df[which(df$sex3=="female"
              #|df$sex3=="female"
              ),]
x$sex3 <- droplevels(x$sex3)

# and remove missing again
sum(is.na(x$moral_identity_symbolic))
x <- x[which(is.na(x$moral_identity_symbolic)==FALSE),]
x <- subset(x, !(country_name %in% exclusions))

x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )
female_means <- lapply(x, means_fun)
female_sds <- lapply(x, sd_fun)

```

### Symbolization Ns

```{r}
#### Total Ns
x <- df[which(df$sex3=="male"
              |df$sex3=="female"
              ),]
x$sex3 <- droplevels(x$sex3)

# and remove missing again
sum(is.na(x$moral_identity_symbolic))
x <- x[which(is.na(x$moral_identity_symbolic)==FALSE),]
x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )
total_Ns <- lapply(x, N_fun)

#### Ns for Males 
# re-load ungrouped data #
x <- df[which(df$sex3=="male"
              #|df$sex3=="female"
              ),]
x$sex3 <- droplevels(x$sex3)

# and remove missing again
sum(is.na(x$moral_identity_symbolic))
x <- x[which(is.na(x$moral_identity_symbolic)==FALSE),]
x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )
male_Ns <- lapply(x, N_fun)

#### Ns for Females
# re-load ungrouped data #
x <- df[which(df$sex3=="female"
              #|df$sex3=="female"
              ),]
x$sex3 <- droplevels(x$sex3)
# and remove missing again
sum(is.na(x$moral_identity_symbolic))
x <- x[which(is.na(x$moral_identity_symbolic)==FALSE),]
x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )
female_Ns <- lapply(x, N_fun)

test <- 
  cbind.data.frame(do.call(rbind, total_Ns)
                 ,do.call(rbind, female_Ns)
                 ,do.call(rbind, male_Ns))

test <- `colnames<-`(test, c("Total N","n Female","N male"))

loop_exclusions_fun <- function(x){
  if(rownames(test)[x] %in% exclusions) print(paste0(rownames(test)[x],"*"),quote = FALSE)
  else print(rownames(test)[x])
  }

rownames(test) <-  unlist(lapply(1:length(rownames(test)), loop_exclusions_fun))
test
#test <-
test[order(rownames(test) ),]

```

```{r include=TRUE}

apa_table(
  test
  , align = c("c", "c", "c")
  , caption = "Total samples for each country for Symbolization analysis"
  , note = "* = sample too small to be included in analysis below")

```

### Symbolization Table

```{r}

test <- 
  cbind.data.frame(
     round(do.call(rbind, female_means),digits=2)
    ,round(do.call(rbind, female_sds),digits=2)
    ,round(do.call(rbind, male_means),digits=2)
    ,round(do.call(rbind, female_sds),digits=2)
    ,round(do.call(rbind, ts_for_table),digits=2)
    ,round(do.call(rbind, df_for_table),digits=2)
    ,do.call(rbind, ps_for_table)
    ,round(do.call(rbind, ds_for_table),digits=2)
    )

test <- `colnames<-`(test, c("mean F","SD F","mean M", "SD M", "t","df","p","d"))
test

x <- df[which(df$sex3=="male"
              |df$sex3=="female"
              ),]
x$sex3 <- droplevels(x$sex3)

# and remove missing again
sum(is.na(x$moral_identity_symbolic))
x <- x[which(is.na(x$moral_identity_symbolic)==FALSE),]
x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )
total_Ns <- lapply(x, N_fun)

P_R <- x$`Puerto Rico`
P_R <- 
  `rownames<-`(
    `colnames<-`(
      as.data.frame(
        rbind(
          # c("mean F","SD F","mean M", "SD M", "t","df","p","d")
          # ,
          c(
            P_R$moral_identity_symbolic[which(P_R$sex3=="male")]
            ,"-"
            ,P_R$moral_identity_symbolic[which(P_R$sex3=="female")]
            ,"-"
            ,"-"
            ,"-"
            ,"-"
            ,"-"))),
      c("mean F","SD F","mean M", "SD M", "t","df","p","d"))
    ,c("Puerto Rico"))

test <- rbind(test[1:49,],P_R,test[50:length(test$`mean F`),])
test <- test[order(rownames(test) ),]

```

```{r include=TRUE}

apa_table(
  test
  , align = c("l", "l", "c", "c", "c", "c", "c", "c", "c", "c")
  , caption = "Gender differences in Symbolization for each Country"
  , note = "* = sig. at p < .05; ** = sig. at p < .001")

```

### Symbolization Plot

```{r symbolizationPlot}

# x <- df[which(df$sex3=="male"|df$sex3=="female"),]
# x$sex3 <- droplevels(x$sex3)
# x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )
# 
# #### identify countries with small samples
# male_length <- function(x){sum(x$sex3=='male')}
# males <- lapply(x, male_length)
# female_length <- function(x){sum(x$sex3=='female')}
# females <- lapply(x, female_length)
# names(females)
# 
# #### generate a vector containing all countries with less than 10 observations for either males or females ####
# exclusions <- c(names(males[males<2]),
#   names(females[females<2]))
# 
# #### re-load ungrouped data ####
# x <- df[which(df$sex3=="male"|df$sex3=="female"),]
# x$sex3 <- droplevels(x$sex3)
# 
# # and remove missing again
# sum(is.na(x$moral_identity_symbolic))
# x <- x[which(is.na(x$moral_identity_symbolic)==FALSE),]
# 
# #### Filter out any countries to be excluded ####
# x <- subset(x, !(country_name %in% exclusions))
# 
# df1 <- x
# 
# df1_male <- df1[which(df1$sex3=="male"),]
# df1_female <- df1[which(df1$sex3=="female"),]
# 
# internal_mean_sd <- function(x){
#   z <- x
#   z$mean <- mean(z$moral_identity_symbolic, na.rm=T)
#   z$sd <- sd(z$moral_identity_symbolic, na.rm=T)
#   z
# }
# 
# 
# df1 <- rbind.data.frame(
#   internal_mean_sd(df1_female),
#   internal_mean_sd(df1_male)
# )
# 
# df1$sex3 <- dplyr::recode(df1$sex3
#                            , "male" = "Male"
#                            , "female" = "Female")
# 
# g1 <- ggplot(df1,aes(x=sex3,y=moral_identity_symbolic))+
#   geom_jitter(height = .1, width = .1, size=.0001, color="gray") +
#   geom_jitter(shape=16, position=position_jitter(height = .5, width = .45), size=.001, color="lightgrey")+
#   geom_violin(fill = NA) +
#   #ggforce::geom_sina(size=.01)+
#   # geom_sina(shape=16
#   #           , position=position_jitter(height=.25,width=0)
#   #           , size=.000001
#   #           , color="dark grey"
#   #           ,jitter_y=F
#   # )+
#   geom_boxplot(width=0.1, outlier.shape = NA, color='grey')+
#   geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.06,
#                 position=position_dodge(0.05), color="#5a5a5a")+
#   stat_summary(fun=mean, geom="point", shape=16, size=.5)+
#   # stat_summary(geom = "linerange",
#   #   fun.data = mean_sdl,
#   #   fun.args = list(mult = 1),
#   #   colour = "black"
#   # )+
#   xlab("Gender")+
#   ylab("Internalization")+
#   facet_wrap(~df1$ISO3, ncol = 9 )+
#   #geom_jitter(shape=16, position=position_jitter(height = .5, width = .45), size=.01, color="dark grey")
#   theme_bw() +
#   theme(panel.border = element_blank(),
#         axis.line = element_line(size = .2),
#         strip.background  = element_blank(),
#         panel.grid = element_blank(),
#         plot.title=element_text(#family="Times",
#           size=12
#         ),
#         #legend.position="right",
#         legend.text=element_text(#family="Times",
#           size=8
#         ),
#         legend.title=element_text(#family="Times",
#           size=10
#         ),
#         axis.text=element_text(#family="Times",
#           colour = "black",
#           size=8
#         ),
#         axis.ticks.x = element_blank(),
#         axis.title=element_text(#family="Times",
#           size=12
#         ),
#         strip.text=element_text(#family = "Times",
#           size = 12
#         ),
#         # strip.background = element_rect(fill = "white"),
#         legend.position="none")

#g1

#ggsave("plots/symbolization.png", g1, width = 40, height = 25, units = "in", dpi = 300)
#g2
```

```{r, SymbolizationGender,fig.cap="Gender differences in Symbolization for each country",include=TRUE}
#suppressWarnings(print(g1))

```

<style>
#wrapper { width: 100%;; height: 702px; padding: 0; overflow: hidden; border: 1px}
#scaled-frame { width: 250%; height: 900px; border: 1px; }
#scaled-frame {
    zoom: 0.50;
    -moz-transform: scale(0.50);
    -moz-transform-origin: 0 0;
    -o-transform: scale(0.50);
    -o-transform-origin: 0 0;
    -webkit-transform: scale(0.50);
    -webkit-transform-origin: 0 0;
}

@media screen and (-webkit-min-device-pixel-ratio:0) {
 #scaled-frame  { zoom: 1;  }
}
</style>

<div id="wrapper"><iframe id="scaled-frame" src="plots/internalization.png"></iframe></div>