### 2023-02-06 supp_H2.Rmd

---
title: "H2 Supplement"
author: "Mariola Paruzel-Czachura, Cillian McHugh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: "resources/bib/My Library.bib"
csl: "resources/bib/apa.csl"
output:
  bookdown::html_document2:
    fig_caption: yes
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 4
    number_sections: false
editor_options: 
  chunk_output_type: console
---

<style>
  .superbigimage{
      overflow-x:scroll;
      white-space: nowrap;
  }

  .superbigimage img{
     max-width: none;
  }

</style>

```{r setuph2supp, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = FALSE)
```

```{r}
rm(list = ls())
library(tidyverse)
library(plyr)
library(psych)
library(skimr)
library(corrplot)
library(Hmisc)
#library(psy)
library(afex)
#devtools::install_github("cillianmiltown/R_desnum")
library(desnum)
library(papaja)
library(purrr)
#install.packages("interactions")
library(interactions)
library(kableExtra)
#devtools::install_github("crsh/papaja")
library("papaja")
library(lme4)
library(car)
library(ggplot2)
library(nlme)
#install.packages("reshape")
library(reshape)
library(corx)
library(lsr)
#install.packages("Cairo")
#library(Cairo)
library(foreign)
library(ltm)
library("ggpubr")
#install.packages("rworldmap")
library(rworldmap)
#devtools::install_github("rensa/ggflags")
library(ggflags)
#devtools::install_github("schliebs/ggoxford")
library(ggoxford)
library(ggforce)
library(ggbeeswarm)
library(countrycode)
```

```{r}

#source("set_up_data_calculate_means.R")
load("data/data_with_means.RData")

df$sex3 <- as.factor(df$sex1)
df$sex3 <- dplyr::recode(df$sex3,
               "1"="male"
              ,"2"="female"
              ,"3"="other")

df$country_lower <- tolower(df$country)
df$country_name <- countrycode(df$ISO3, origin = "iso3c", destination = "country.name")
```

```{r}
rm(list = ls())
#source("set_up_data_calculate_means.R")
load("data/data_with_means.RData")

df$sex3 <- as.factor(df$sex1)
df$sex3 <- dplyr::recode(df$sex3,
               "1"="male"
              ,"2"="female"
              ,"3"="other")

df$country_lower <- tolower(df$country)
df$country_name <- countrycode(df$ISO3, origin = "iso3c", destination = "country.name")
```

## Age & Internalization

```{r}

ps <- function(y){
  y <- round(y, digits = 3)
  if(as.numeric(sqrt( y*y) ) <.001) print(paste0("<", " ", ".001","**"), quote = FALSE)
  else if(as.numeric(sqrt( y*y) ) <.05) print(paste0(sub("^(-?)0.", "\\1.", sprintf("%.3f", 
        y, quote = FALSE)),"*"))
  else print(sub("^(-?)0.", "\\1.", sprintf("%.3f", 
        y)))}

x <- df
x <- x[which(is.na(x$moral_identity_internal)==FALSE),]
x <- x[which(is.na(x$age)==FALSE),]
#### group by country for analysis ####
x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )

N_fun <- function(x){
  length(x$sex3)
}
total_Ns <- lapply(x, N_fun)

exclusions <- names(total_Ns)[which(total_Ns<3)]
x <- df
x <- x[which(is.na(x$moral_identity_internal)==FALSE),]
x <- x[which(is.na(x$age)==FALSE),]
x <- subset(x, !(country_name %in% exclusions))

x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )

cor_fun <- function(x){
  cor.test(x$moral_identity_symbolic,x$age)
}
all_cor_tests <- lapply(x, cor_fun)

all_cor_rs <- function(c){
  c$estimate
}
rs_for_table <- lapply(all_cor_tests, all_cor_rs)
all_cor_ps <- function(c){
  c$estimate
}
ps_for_table <- lapply(all_cor_tests, all_cor_ps)
ps_output_for_table <- lapply((lapply(all_cor_tests, all_cor_ps)), ps)
Ns_for_table <- lapply(x, N_fun)

means_INT_fun <- function(x){
  mean(x$moral_identity_internal)
}
mean_int_for_table <- lapply(x, means_INT_fun)
sd_INT_fun <- function(x){
  sd(x$moral_identity_internal)
}
sd_int_for_table <- lapply(x, sd_INT_fun)

means_age_fun <- function(x){
  mean(x$age)
}
mean_age_for_table <- lapply(x, means_age_fun)
sd_age_fun <- function(x){
  sd(x$age)
}
sd_age_for_table <- lapply(x, sd_age_fun)

```

```{r}

test <- 
  cbind.data.frame(
     round(do.call(rbind,mean_age_for_table),digits = 2)
    ,round(do.call(rbind,sd_age_for_table),digits = 2)
    ,round(do.call(rbind,mean_int_for_table),digits = 2)
    ,round(do.call(rbind,sd_int_for_table),digits = 2)
    ,round(do.call(rbind, Ns_for_table))
    ,round(do.call(rbind, rs_for_table),digits=3)
    ,do.call(rbind, ps_output_for_table)
    )

test <- `colnames<-`(test, c("mean age","SD age","mean internalization", "SD internalization","N", "r","p"))
test

sum(unlist(ps_for_table)<.05)

test1 <- subset(test, p < .05)
test1

total_sig <- length(test1$`mean age`)

neg <- sum(test1$r <0 )
pos <- sum(test1$r>0)

x <- df
x <- x[which(is.na(x$moral_identity_internal)==FALSE),]
x <- x[which(is.na(x$age)==FALSE),]

x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )

P_R <- x$`Puerto Rico`
P_R <- 
  `rownames<-`(
    `colnames<-`(
      as.data.frame(
        rbind(
          # c("mean F","SD F","mean M", "SD M", "t","df","p","d")
          # ,
          c(
            round(mean(P_R$age),digits=2)
            ,round(sd(P_R$age),digits=2)
            ,round(mean(P_R$moral_identity_internal),digits=2)
            ,round(sd(P_R$moral_identity_internal),digits=2)
            ,length(P_R$moral_identity_internal)
            ,"-"
            ,"-"))),
      c("mean age","SD age","mean internalization", "SD internalization","N", "r","p"))
    ,c("Puerto Rico"))

test <- rbind(test[1:49,],P_R,test[50:length(test$p),])
test <- test[order(rownames(test) ),]

```

```{r include=TRUE}

apa_table(
  test
  , align = c("l", "l", "c", "c", "c", "c", "c", "c", "c", "c")
  , caption = "Correlation between Age and Internalization for each Country"
  , note = "* = sig. at p < .05; ** = sig. at p < .001")

```

```{r}

test <- 
  cbind.data.frame(
     round(do.call(rbind,mean_age_for_table),digits = 2)
    ,round(do.call(rbind,sd_age_for_table),digits = 2)
    ,round(do.call(rbind,mean_int_for_table),digits = 2)
    ,round(do.call(rbind,sd_int_for_table),digits = 2)
    ,round(do.call(rbind, Ns_for_table))
    ,round(do.call(rbind, rs_for_table),digits=3)
    ,do.call(rbind, ps_for_table)
    )

test <- `colnames<-`(test, c("mean age","SD age","mean internalization", "SD internalization","N", "r","p"))
test

sig_countries <- rownames(test[which(test$p<.05),])

# loop_sigs_fun <- function(x){
#   if(rownames(test)[x] %in% sig_countries) print(paste0(rownames(test)[x],"*"),quote = FALSE)
#   else print(rownames(test)[x])
# }
# 
# rownames(test) <-  unlist(lapply(1:length(rownames(test)), loop_sigs_fun))
# 
# loop_sigs_fun2 <- function(x,y){
#   if(x$country_name[y] %in% sig_countries) print(paste0(x$country_name[y],"*"),quote = FALSE)
#   else print(x$country_name[y])
# }

```

```{r}

df1 <- subset(df, !(country_name %in% exclusions))

loop_sigs_fun3 <- function(y){
  if(df1$country_name[y] %in% sig_countries) print(paste0(df1$country_name[y],"*"),quote = FALSE)
  else print(df1$country_name[y])
}

df1$country_name2 <- unlist(lapply(1:length(df1$country_name),loop_sigs_fun3))

N_fun <- function(x){
  length(x$sex3)
}
length(df1$sex3)
x <- df1
x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )
df1 <- left_join(df1
                      ,  
                      `colnames<-`(
                        cbind.data.frame(
                          unlist(lapply(x, N_fun))
                          ,names(unlist(lapply(x, N_fun))))
                        ,c("N","country_name"))
                      , by = "country_name")

df1$country_name3 <- print(paste0(df1$country_name2, " (N = ", df1$N,")"))

g1 <- ggplot(df1, aes(age, moral_identity_internal))+
  geom_jitter(size=.05, color='darkgrey') + 
  ylim(0,10) +
  geom_smooth(method=lm, size = .5,color='black') +
  labs(y = "Internalization",
       x = "Age"
  )+ 
  theme_apa()+
  theme_bw()+
  theme(panel.border = element_blank(),
        axis.line = element_line(size = .2),
        strip.background  = element_blank(),
        #panel.grid = element_blank(),
        plot.title=element_text(#family="Times",
          size=12
        ),
        #legend.position="right",
        legend.text=element_text(#family="Times",
          size=8
        ),
        legend.title=element_text(#family="Times",
          size=10
        ),
        axis.text=element_text(#family="Times",
          colour = "black",
          size=8
        ),
        axis.ticks.x = element_blank(),
        axis.title=element_text(#family="Times",
          size=12
        ),
        strip.text=element_text(#family = "Times",
          size = 12
        ),
        # strip.background = element_rect(fill = "white"),
        legend.position="none")+
  facet_wrap(~df1$country_name3, ncol = 9 )

ggsave("plots/age_internalization.png", g1, width = 40, height = 25, units = "in", dpi = 300)

```

(<a href="https://raw.githubusercontent.com/cillianmiltown/moral_ID_gender/main/plots/age_internalization.png" target="_blank">open in new window</a>)

<style>
#wrapper { width: 100%;; height: 702px; padding: 0; overflow: hidden; border: 1px}
#scaled-frame { width: 250%; height: 900px; border: 1px; }
#scaled-frame {
    zoom: 0.50;
    -moz-transform: scale(0.50);
    -moz-transform-origin: 0 0;
    -o-transform: scale(0.50);
    -o-transform-origin: 0 0;
    -webkit-transform: scale(0.50);
    -webkit-transform-origin: 0 0;
}

@media screen and (-webkit-min-device-pixel-ratio:0) {
 #scaled-frame  { zoom: 1;  }
}
</style>

<div id="wrapper"><iframe id="scaled-frame" src="plots/age_internalization.png"></iframe></div>

## Age & Symbolization

```{r}

ps <- function(y){
  y <- round(y, digits = 3)
  if(as.numeric(sqrt( y*y) ) <.001) print(paste0("<", " ", ".001","**"), quote = FALSE)
  else if(as.numeric(sqrt( y*y) ) <.05) print(paste0(sub("^(-?)0.", "\\1.", sprintf("%.3f", 
        y, quote = FALSE)),"*"))
  else print(sub("^(-?)0.", "\\1.", sprintf("%.3f", 
        y)))}

x <- df
x <- x[which(is.na(x$moral_identity_symbolic)==FALSE),]
x <- x[which(is.na(x$age)==FALSE),]
#### group by country for analysis ####
x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )

N_fun <- function(x){
  length(x$sex3)
}
total_Ns <- lapply(x, N_fun)

exclusions <- names(total_Ns)[which(total_Ns<3)]
x <- df
x <- x[which(is.na(x$moral_identity_symbolic)==FALSE),]
x <- x[which(is.na(x$age)==FALSE),]
x <- subset(x, !(country_name %in% exclusions))

x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )

cor_fun <- function(x){
  cor.test(x$moral_identity_symbolic,x$age)
}
all_cor_tests <- lapply(x, cor_fun)

all_cor_rs <- function(c){
  c$estimate
}
rs_for_table <- lapply(all_cor_tests, all_cor_rs)
all_cor_ps <- function(c){
  c$estimate
}
ps_for_table <- lapply(all_cor_tests, all_cor_ps)
ps_output_for_table <- lapply((lapply(all_cor_tests, all_cor_ps)), ps)
Ns_for_table <- lapply(x, N_fun)

means_INT_fun <- function(x){
  mean(x$moral_identity_symbolic)
}
mean_int_for_table <- lapply(x, means_INT_fun)
sd_INT_fun <- function(x){
  sd(x$moral_identity_symbolic)
}
sd_int_for_table <- lapply(x, sd_INT_fun)

means_age_fun <- function(x){
  mean(x$age)
}
mean_age_for_table <- lapply(x, means_age_fun)
sd_age_fun <- function(x){
  sd(x$age)
}
sd_age_for_table <- lapply(x, sd_age_fun)

```

```{r}

test <- 
  cbind.data.frame(
     round(do.call(rbind,mean_age_for_table),digits = 2)
    ,round(do.call(rbind,sd_age_for_table),digits = 2)
    ,round(do.call(rbind,mean_int_for_table),digits = 2)
    ,round(do.call(rbind,sd_int_for_table),digits = 2)
    ,round(do.call(rbind, Ns_for_table))
    ,round(do.call(rbind, rs_for_table),digits=3)
    ,do.call(rbind, ps_output_for_table)
    )

test <- `colnames<-`(test, c("mean age","SD age","mean symbolization", "SD symbolization","N", "r","p"))
test

sum(unlist(ps_for_table)<.05)

test1 <- subset(test, p < .05)
test1

total_sig <- length(test1$`mean age`)

neg <- sum(test1$r <0 )
pos <- sum(test1$r>0)

x <- df
x <- x[which(is.na(x$moral_identity_symbolic)==FALSE),]
x <- x[which(is.na(x$age)==FALSE),]

x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )

P_R <- x$`Puerto Rico`
P_R <- 
  `rownames<-`(
    `colnames<-`(
      as.data.frame(
        rbind(
          # c("mean F","SD F","mean M", "SD M", "t","df","p","d")
          # ,
          c(
             round(mean(P_R$age),digits=2)
            ,round(sd(P_R$age),digits=2)
            ,round(mean(P_R$moral_identity_symbolic),digits=2)
            ,round(sd(P_R$moral_identity_symbolic),digits=2)
            ,length(P_R$moral_identity_symbolic)
            ,"-"
            ,"-"))),
      c("mean age","SD age","mean symbolization", "SD symbolization","N", "r","p"))
    ,c("Puerto Rico"))

test <- rbind(test[1:49,],P_R,test[50:length(test$p),])
test <- test[order(rownames(test) ),]

```

```{r include=TRUE}

apa_table(
  test
  , align = c("l", "l", "c", "c", "c", "c", "c", "c", "c", "c")
  , caption = "Correlation between Age and Symbolization for each Country"
  , note = "* = sig. at p < .05; ** = sig. at p < .001")

```

```{r}

test <- 
  cbind.data.frame(
     round(do.call(rbind,mean_age_for_table),digits = 2)
    ,round(do.call(rbind,sd_age_for_table),digits = 2)
    ,round(do.call(rbind,mean_int_for_table),digits = 2)
    ,round(do.call(rbind,sd_int_for_table),digits = 2)
    ,round(do.call(rbind, Ns_for_table))
    ,round(do.call(rbind, rs_for_table),digits=3)
    ,do.call(rbind, ps_for_table)
    )

test <- `colnames<-`(test, c("mean age","SD age","mean symbolization", "SD internalization","N", "r","p"))
test

sig_countries <- rownames(test[which(test$p<.05),])

# loop_sigs_fun <- function(x){
#   if(rownames(test)[x] %in% sig_countries) print(paste0(rownames(test)[x],"*"),quote = FALSE)
#   else print(rownames(test)[x])
# }
# 
# rownames(test) <-  unlist(lapply(1:length(rownames(test)), loop_sigs_fun))
# 
# loop_sigs_fun2 <- function(x,y){
#   if(x$country_name[y] %in% sig_countries) print(paste0(x$country_name[y],"*"),quote = FALSE)
#   else print(x$country_name[y])
# }

```

```{r}

df1 <- subset(df, !(country_name %in% exclusions))

loop_sigs_fun3 <- function(y){
  if(df1$country_name[y] %in% sig_countries) print(paste0(df1$country_name[y],"*"),quote = FALSE)
  else print(df1$country_name[y])
}

df1$country_name2 <- unlist(lapply(1:length(df1$country_name),loop_sigs_fun3))

N_fun <- function(x){
  length(x$sex3)
}
length(df1$sex3)
x <- df1
x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )
df1 <- left_join(df1
                      ,  
                      `colnames<-`(
                        cbind.data.frame(
                          unlist(lapply(x, N_fun))
                          ,names(unlist(lapply(x, N_fun))))
                        ,c("N","country_name"))
                      , by = "country_name")

df1$country_name3 <- print(paste0(df1$country_name2, " (N = ", df1$N,")"))

g1 <- ggplot(df1, aes(age, moral_identity_symbolic))+
  geom_jitter(size=.05, color='darkgrey') + 
  ylim(0,10) +
  geom_smooth(method=lm, size = .5,color='black') +
  labs(y = "Symbolization",
       x = "Age"
  )+ 
  theme_apa()+
  theme_bw()+
  theme(panel.border = element_blank(),
        axis.line = element_line(size = .2),
        strip.background  = element_blank(),
        #panel.grid = element_blank(),
        plot.title=element_text(#family="Times",
          size=12
        ),
        #legend.position="right",
        legend.text=element_text(#family="Times",
          size=8
        ),
        legend.title=element_text(#family="Times",
          size=10
        ),
        axis.text=element_text(#family="Times",
          colour = "black",
          size=8
        ),
        axis.ticks.x = element_blank(),
        axis.title=element_text(#family="Times",
          size=12
        ),
        strip.text=element_text(#family = "Times",
          size = 12
        ),
        # strip.background = element_rect(fill = "white"),
        legend.position="none")+
  facet_wrap(~df1$country_name3, ncol = 9 )

ggsave("plots/age_symbolization.png", g1, width = 40, height = 25, units = "in", dpi = 300)

```

(<a href="https://raw.githubusercontent.com/cillianmiltown/moral_ID_gender/main/plots/age_symbolization.png" target="_blank">open in new window</a>)

<style>
#wrapper { width: 100%;; height: 702px; padding: 0; overflow: hidden; border: 1px}
#scaled-frame { width: 250%; height: 900px; border: 1px; }
#scaled-frame {
    zoom: 0.50;
    -moz-transform: scale(0.50);
    -moz-transform-origin: 0 0;
    -o-transform: scale(0.50);
    -o-transform-origin: 0 0;
    -webkit-transform: scale(0.50);
    -webkit-transform-origin: 0 0;
}

@media screen and (-webkit-min-device-pixel-ratio:0) {
 #scaled-frame  { zoom: 1;  }
}
</style>

<div id="wrapper"><iframe id="scaled-frame" src="plots/age_symbolization.png"></iframe></div>