### 2023-02-10 H3.Rmd

---
title: "H3: Cultural Differences in Symbolization"
author: "Mariola Paruzel-Czachura, Cillian McHugh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: "resources/bib/My Library.bib"
csl: "resources/bib/apa.csl"
output:
  bookdown::html_document2:
    fig_caption: yes
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 4
    number_sections: false
editor_options: 
  chunk_output_type: console
---

<style>
  .superbigimage{
      overflow-x:scroll;
      white-space: nowrap;
  }

  .superbigimage img{
     max-width: none;
  }

</style>

```{r setuph3, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = FALSE)
```

```{r}
rm(list = ls())
library(tidyverse)
library(plyr)
library(psych)
library(skimr)
library(corrplot)
library(Hmisc)
#library(psy)
library(afex)
#devtools::install_github("cillianmiltown/R_desnum")
library(desnum)
library(papaja)
library(purrr)
#install.packages("interactions")
library(interactions)
library(kableExtra)
#devtools::install_github("crsh/papaja")
library("papaja")
library(lme4)
library(car)
library(ggplot2)
library(nlme)
#install.packages("reshape")
library(reshape)
library(corx)
library(lsr)
#install.packages("Cairo")
#library(Cairo)
library(foreign)
library(ltm)
library("ggpubr")
#install.packages("rworldmap")
library(rworldmap)
#devtools::install_github("rensa/ggflags")
library(ggflags)
#devtools::install_github("schliebs/ggoxford")
library(ggoxford)
library(ggforce)
library(ggbeeswarm)
library(countrycode)

```

```{r}
rm(list = ls())

#source("set_up_data_calculate_means.R")
load("data/data_with_means.RData")

df$sex3 <- as.factor(df$sex1)
df$sex3 <- dplyr::recode(df$sex3,
               "1"="male"
              ,"2"="female"
              ,"3"="other")

df$country_lower <- tolower(df$country)
df$country_name <- countrycode(df$ISO3, origin = "iso3c", destination = "country.name")

x <- df
```

```{r}

# h3a yes
# h3b n/a (competing)
# h3c yes
# h3d no (no relationship)
# h3e yes
# h3f yes
# h3g no, no, yes

x <- x[which(is.na(x$moral_identity_symbolic)==FALSE),]

fit <- lm(moral_identity_symbolic ~
                    # sex3
                    # + age
                     Power.Distance
                    + Individualism
                    + Masculinity
                    + Uncertainty.Avoidance
                    # + Long.Term.Orientation
                    # + Indulgence
                    # + CTL_DG
                    # + CTL_DS
                    # + CTL_C
                , data = x
            )
fit
fit_results <- summary(fit)
fit_results

fit_results$adj.r.squared
anova(fit)
apa_fit_results <- apa_print(fit)

cor.test(x$Power.Distance,x$moral_identity_symbolic)

apa_fit_results$full_result$modelfit$r2
fit$qr
apa_fit_results$table
results_for_table <- apa_fit_results$table

estimate_for_table <-
  `colnames<-`(
    cbind.data.frame(
      unlist(fit_results$coefficients[,1])
      ,names(unlist(fit_results$coefficients[,1]))
    ), c("estimate","rows"))$estimate

stderror_for_table <-
  `colnames<-`(
    cbind.data.frame(
      unlist(fit_results$coefficients[,2])
      ,names(unlist(fit_results$coefficients[,2]))
    ), c("stderror","rows"))$stderror

estimate_for_table
stderror_for_table

results_for_table$estimate <- estimate_for_table
results_for_table$StdError <- stderror_for_table

results_for_table <- results_for_table[c(1,2,7,3,4,5,6)]

results_for_table$estimate <- sprintf("%.3f", results_for_table$estimate)
results_for_table$StdError <- sprintf("%.3f", results_for_table$StdError)

# library(Hmisc)
# 
# var.labels = 
#   c(
#     term        = "Predictor" 
#     , "estimate"  = "$b$" 
#     , "StdError"  = "$S.E.$"
#     , conf.int  = "95\\% CI" 
#     , statistic = "$t$" 
#     , df        = "$\\mathit{df}$" 
#     , p.value   = "$p$" 
#     )
# 
# label(results_for_table) = as.list(var.labels[match(names(results_for_table), names(var.labels))])

results_for_table$term <- c(
  "Intercept"
  ,"Power Distance"
  ,"Individualism"
  ,"Masculinity"
  ,"Uncertainty Avoidance"
  ,"Long Term Orientation"
  ,"Indulgence"
  ,"Tightness: Domain General"
  ,"Tightness: Domain Specific"
  ,"Tightness: Combination Measure"
)

variable_label(results_for_table$estimate) <- "$b$"
variable_label(results_for_table$StdError) <- "$S.E.$"
variable_label(results_for_table$term) <- "Predictor"

```

```{r}

# h3a yes
# h3b n/a (competing)
# h3c yes
# h3d no (no relationship)
# h3e yes
# h3f yes
# h3g no, no, yes

x <- x[which(is.na(x$moral_identity_symbolic)==FALSE),]

fit <- lm(moral_identity_symbolic ~
                    # sex3
                    # + age
                     Power.Distance
                    + Individualism
                    + Masculinity
                    + Uncertainty.Avoidance
                    + Long.Term.Orientation
                    + Indulgence
                    + CTL_DG
                    + CTL_DS
                    + CTL_C
                , data = x
            )

fit_results <- summary(fit)
fit_results$adj.r.squared
anova(fit)
apa_fit_results <- apa_print(fit)

cor.test(x$Power.Distance,x$moral_identity_symbolic)

apa_fit_results$full_result$modelfit$r2
fit$qr
apa_fit_results$table
results_for_table <- apa_fit_results$table

estimate_for_table <-
  `colnames<-`(
    cbind.data.frame(
      unlist(fit_results$coefficients[,1])
      ,names(unlist(fit_results$coefficients[,1]))
    ), c("estimate","rows"))$estimate

stderror_for_table <-
  `colnames<-`(
    cbind.data.frame(
      unlist(fit_results$coefficients[,2])
      ,names(unlist(fit_results$coefficients[,2]))
    ), c("stderror","rows"))$stderror

estimate_for_table
stderror_for_table

results_for_table$estimate <- estimate_for_table
results_for_table$StdError <- stderror_for_table

results_for_table <- results_for_table[c(1,2,7,3,4,5,6)]

results_for_table$estimate <- sprintf("%.3f", results_for_table$estimate)
results_for_table$StdError <- sprintf("%.3f", results_for_table$StdError)

# library(Hmisc)
# 
# var.labels = 
#   c(
#     term        = "Predictor" 
#     , "estimate"  = "$b$" 
#     , "StdError"  = "$S.E.$"
#     , conf.int  = "95\\% CI" 
#     , statistic = "$t$" 
#     , df        = "$\\mathit{df}$" 
#     , p.value   = "$p$" 
#     )
# 
# label(results_for_table) = as.list(var.labels[match(names(results_for_table), names(var.labels))])

results_for_table$term <- c(
  "Intercept"
  ,"Power Distance"
  ,"Individualism"
  ,"Masculinity"
  ,"Uncertainty Avoidance"
  ,"Long Term Orientation"
  ,"Indulgence"
  ,"Tightness: Domain General"
  ,"Tightness: Domain Specific"
  ,"Tightness: Combination Measure"
)

variable_label(results_for_table$estimate) <- "$b$"
variable_label(results_for_table$StdError) <- "$S.E.$"
variable_label(results_for_table$term) <- "Predictor"

```

Our third hypothesis predicted relationships between symbolization and specific cultural variables. The variables question are: power distance, individualism, masculinity, uncertainty avoidance, long term orientation, and indulgence, taken from Hofstede (2001), as well as three measures of cultural tightness (domain-specific, domain-general, and a combination index). These variables were inputted at the country level, and as such, we conducted a hierarchical linear regression model to investigate relationship with symbolization. Symbolization was our outcome measure and each cultural variable was included as a predictor in the model.

Overall the model was a significant predictor of symbolization `r apa_fit_results$full_result$modelfit$r2`. The full regression results are displayed in Table\ \@ref(tab:regressionTable).

```{r regressionTable, include=TRUE}
apa_table(
  results_for_table
#  , align = c("c", "c", "c")
  , caption = "Full Regression Results for each country level variable predicting symbolization"
  )

```

In line with H3a greater power distance was associated with higher symbolization scores.

We proposed two competing hypotheses for the relationship between symbolization and symbolization. H3b~1~ predicted that symbolization would be negatively related to individualism (i.e., higher symbolization would be observed in more collectivist cultures). H3b~2~ predicted that individualism and symbolization would be positively associated. Our results support H3b~2~ (over H3b~1~), suggesting that members of more individualistic societies show greater symbolization scores.

We predicted that members of more masculine cultures would score higher in symbolization (H3c). Our results were in line with this prediction.

For H3d we predicted that uncertainty avoidance would be related to higher symbolization. This was not supported, and we found no relationship between symbolization and uncertainty avoidance.

We predicted a negative association between long-term orientation and symbolization (H3e), that is, we expected participants in societies showing greater long-term orientation to present with lower symbolization scores. Our results supported this hypothesis.

Regarding indulgence/restraint, we predicted (H3f) that greater indulgence would be related to higher symbolization scores, and this hypothesis was supported.

We employed measures of three dimensions of cultural tightness, domain-general, domain-specific, and combination, and predicted that symbolization would be positively associated with each dimension. Our hypothesis was supported for only one dimension (domain-general). Interestingly, both domain-specific and combination dimensions of cultural tightness were negatively related to symbolization.