### 2023-07-31 H2.Rmd
added list of countries for age correlations

---
title: "H2: Age Differences"
author: "Mariola Paruzel-Czachura, Cillian McHugh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: "resources/bib/My Library.bib"
csl: "resources/bib/apa.csl"
output:
  bookdown::html_document2:
    fig_caption: yes
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 4
    number_sections: false
editor_options: 
  chunk_output_type: console
---

<style>
  .superbigimage{
      overflow-x:scroll;
      white-space: nowrap;
  }

  .superbigimage img{
     max-width: none;
  }

</style>

```{r setuph2, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = FALSE)
```

```{r}
rm(list = ls())
library(tidyverse)
library(plyr)
library(psych)
library(skimr)
library(corrplot)
library(Hmisc)
#library(psy)
library(afex)
#devtools::install_github("cillianmiltown/R_desnum")
library(desnum)
library(papaja)
library(purrr)
#install.packages("interactions")
library(interactions)
library(kableExtra)
#devtools::install_github("crsh/papaja")
library("papaja")
library(lme4)
library(car)
library(ggplot2)
library(nlme)
#install.packages("reshape")
library(reshape)
library(corx)
library(lsr)
#install.packages("Cairo")
#library(Cairo)
library(foreign)
library(ltm)
library("ggpubr")
#install.packages("rworldmap")
library(rworldmap)
#devtools::install_github("rensa/ggflags")
library(ggflags)
#devtools::install_github("schliebs/ggoxford")
library(ggoxford)
library(ggforce)
library(ggbeeswarm)
library(countrycode)
```

```{r}
rm(list = ls())

#source("set_up_data_calculate_means.R")
load("data/data_with_means.RData")

df$sex3 <- as.factor(df$sex1)
df$sex3 <- dplyr::recode(df$sex3,
               "1"="male"
              ,"2"="female"
              ,"3"="other")

df$country_lower <- tolower(df$country)
df$country_name <- countrycode(df$ISO3, origin = "iso3c", destination = "country.name")

```

```{r}

x <- df
x <- x[which(is.na(x$moral_identity_internal)==FALSE),]
x <- x[which(is.na(x$age)==FALSE),]

c1 <- cor.test(x$moral_identity_internal,x$age)

```

```{r}

x <- df
x <- x[which(is.na(x$moral_identity_symbolic)==FALSE),]
x <- x[which(is.na(x$age)==FALSE),]

c2 <- cor.test(x$moral_identity_symbolic,x$age)

```

Due to mixed findings in the literature, we did not have specific predictions regarding the relationship between age and moral identity, and the analyses reported here are exploratory. Overall there was a significant negative correlation between age and internalization *r* = `r sprintf("%.3f", c1$estimate)`, *p* = `r paste(p_report(c1$p.value))`, suggesting that older participants had lower internalization scores. Similarly, we found a significant negative correlation between age and symbolization *r* = `r sprintf("%.3f", c2$estimate)`, *p* = `r paste(p_report(c2$p.value))`, again, suggesting that older participants had lower symbolization scores than younger participants.

```{r}
x <- df
x <- x[which(is.na(x$moral_identity_internal)==FALSE),]
x <- x[which(is.na(x$age)==FALSE),]

model0 <- lmerTest::lmer(moral_identity_internal ~
                 #  sex_01
                 # * age
                  + (1|country)
                , data = x
                # , contrasts = list(sex3 = contr.sum, valence = contr.sum )
            )

model1 <- lmerTest::lmer(moral_identity_internal ~
                 #  sex_01
                   age
                  + (1|country)
                , data = x
               # , contrasts = list(sex_01 = contr.sum)
            )
summary(model1)
anova(model0,model1)

summary(model1)
anova(model1)
results_anova <- as.data.frame(anova(model0,model1))
results_anova
results_anova$Chisq[2]
results_anova$Df
p1 <- results_anova$`Pr(>Chisq)`[2]
p_report(p1)
#p_report(results_anova$`Pr(>Chisq)`)[2]
results_anova$AIC
summary_model1 <- summary(model1)
summary_model1
results_coef <- as.data.frame(summary_model1$coefficients)
results_coef

aov1 <- anova(model1)

f3a <- aov1$`F value`[1]
p3a <- aov1$`Pr(>F)`[1]

results_coef$Estimate[2]
results_coef$`Std. Error`[2]
results_coef$df[2]
t1 <- results_coef$`t value`[2]
p2 <- results_coef$`Pr(>|t|)`[2]

#QuantPsyc::lm.beta(results_coef)
anova(model1)

```

```{r}

x <- df
x <- x[which(is.na(x$moral_identity_internal)==FALSE),]
x <- x[which(is.na(x$age)==FALSE),]
#### group by country for analysis ####
x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )

N_fun <- function(x){
  length(x$sex3)
}
total_Ns <- lapply(x, N_fun)

exclusions <- names(total_Ns)[which(total_Ns<3)]
x <- df
x <- x[which(is.na(x$moral_identity_internal)==FALSE),]
x <- x[which(is.na(x$age)==FALSE),]
x <- subset(x, !(country_name %in% exclusions))

x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )

cor_fun <- function(x){
  cor.test(x$moral_identity_symbolic,x$age)
}
all_cor_tests <- lapply(x, cor_fun)

all_cor_rs <- function(c){
  c$estimate
}
rs_for_table <- lapply(all_cor_tests, all_cor_rs)
all_cor_ps <- function(c){
  c$p.value
}
ps_for_table <- lapply(all_cor_tests, all_cor_ps)
Ns_for_table <- lapply(x, N_fun)

means_INT_fun <- function(x){
  mean(x$moral_identity_internal)
}
mean_int_for_table <- lapply(x, means_INT_fun)
sd_INT_fun <- function(x){
  sd(x$moral_identity_internal)
}
sd_int_for_table <- lapply(x, sd_INT_fun)

means_age_fun <- function(x){
  mean(x$age)
}
mean_age_for_table <- lapply(x, means_age_fun)
sd_age_fun <- function(x){
  sd(x$age)
}
sd_age_for_table <- lapply(x, sd_age_fun)

```

```{r}
test <- 
  cbind.data.frame(
     do.call(rbind,mean_age_for_table)
    ,do.call(rbind,sd_age_for_table)
    ,do.call(rbind,mean_int_for_table)
    ,do.call(rbind,sd_int_for_table)
    ,round(do.call(rbind, rs_for_table),digits=3)
    ,do.call(rbind, ps_for_table)
    ,round(do.call(rbind, Ns_for_table))
    )

test <- `colnames<-`(test, c("mean age","SD age","mean internalization", "SD internalization", "r","p","N"))
test

sum(unlist(ps_for_table)<.05)

test1 <- subset(test, p < .05)
test1

rownames(subset(test, p < .05))

total_sig <- length(test1$`mean age`)

neg <- sum(test1$r <0 )
pos <- sum(test1$r>0)
neg

noquote(paste0(rownames(test1[which(test1$r<0),]),sep=", "))
noquote(paste0(rownames(test1[which(test1$r>0),]),sep=", "))

neg_names <- rownames(test1[which(test1$r<0),])
neg_names_first <- c(paste0(neg_names[1:(length(neg_names)-1)], sep = ", "))
neg_names_end <- c("and " , neg_names[length(neg_names)])
noquote(c(neg_names_first, neg_names_end))

pos_names <- rownames(test1[which(test1$r>0),])
pos_names_first <- c(paste0(pos_names[1:(length(pos_names)-1)], sep = ", "))
pos_names_end <- c("and " , pos_names[length(pos_names)])
noquote(c(pos_names_first, pos_names_end))

test1$r>0

```

We conducted a linear-mixed-effects model to test if this relationship between age and internalization holds when participants are nested within country. Our outcome measure was internalization and our predictor variable was age; we allowed intercepts to vary by country.
Overall, the model significantly predicted internalization, and provided a better fit for the data than the baseline model,
$\chi$^2^(`r results_anova$Df[2]`) = `r sprintf("%.2f",results_anova$Chisq[2])`, *p* `r paste(p_report(p1))`. 
There was a significant main effect for age,
*F*(`r aov1$NumDF[1]`, `r sprintf("%.2f",round(aov1$DenDF[1],digits=2))`) = 
`r sprintf("%.2f",f3a)`, *p* `r paste(p_report(p3a))`;
with a significant positive association between age and internalization scores, $b$ = `r sprintf("%.3f",results_coef$Estimate[2])`,
*t*(`r sprintf("%.2f",results_coef$df[2])`) = 
`r sprintf("%.2f",t1)`,
*p* `r paste(p_report(p2))`.
To interpret this we conducted a series of bivariate correlations to test if age was related to internalization across all countries.  There was a significant correlation between age and internalization in `r total_sig` countries. Of these, there was a positive correlation between age and internalization in `r pos` countries, and a negative correlation in `r neg` countries. The countries with a positive correlation were `r noquote(paste0(c(pos_names_first, pos_names_end), collapse = ""))`, and the countries with a negative correlation were `r noquote(paste0(c(neg_names_first, neg_names_end), collapse = ""))`  (see Supplementary Materials [Table S_X](https://cillianmiltown.github.io/moral_ID_gender/drafting_supplementary_analysis.html#Age__Internalization)).

```{r}

x <- df
x <- x[which(is.na(x$moral_identity_symbolic)==FALSE),]
x <- x[which(is.na(x$age)==FALSE),]

model0 <- lmerTest::lmer(moral_identity_symbolic ~
                 #  sex_01
                 # * age
                  + (1|country)
                , data = x
                # , contrasts = list(sex3 = contr.sum, valence = contr.sum )
            )

model1 <- lmerTest::lmer(moral_identity_symbolic ~
                 #  sex_01
                   age
                  + (1|country)
                , data = x
               # , contrasts = list(sex_01 = contr.sum)
            )
summary(model1)
anova(model0,model1)

summary(model1)
anova(model1)
results_anova <- as.data.frame(anova(model0,model1))
results_anova
results_anova$Chisq[2]
results_anova$Df
p1 <- results_anova$`Pr(>Chisq)`[2]
p_report(p1)
#p_report(results_anova$`Pr(>Chisq)`)[2]
results_anova$AIC
summary_model1 <- summary(model1)
summary_model1
results_coef <- as.data.frame(summary_model1$coefficients)
results_coef

aov1 <- anova(model1)

f3a <- aov1$`F value`[1]
p3a <- aov1$`Pr(>F)`[1]

results_coef$Estimate[2]
results_coef$`Std. Error`[2]
results_coef$df[2]
t1 <- results_coef$`t value`[2]
p2 <- results_coef$`Pr(>|t|)`[2]

#QuantPsyc::lm.beta(results_coef)
anova(model1)

```

```{r}

x <- df
x <- x[which(is.na(x$moral_identity_symbolic)==FALSE),]
x <- x[which(is.na(x$age)==FALSE),]
#### group by country for analysis ####
x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )

N_fun <- function(x){
  length(x$sex3)
}
total_Ns <- lapply(x, N_fun)

exclusions <- names(total_Ns)[which(total_Ns<3)]
x <- df
x <- x[which(is.na(x$moral_identity_symbolic)==FALSE),]
x <- x[which(is.na(x$age)==FALSE),]
x <- subset(x, !(country_name %in% exclusions))

x <- x %>% group_by(country_name) %>% do(data = (.)) %>% with( set_names(data, country_name) )

cor_fun <- function(x){
  cor.test(x$moral_identity_symbolic,x$age)
}
all_cor_tests <- lapply(x, cor_fun)

all_cor_rs <- function(c){
  c$estimate
}
rs_for_table <- lapply(all_cor_tests, all_cor_rs)
all_cor_ps <- function(c){
  c$p.value
}
ps_for_table <- lapply(all_cor_tests, all_cor_ps)
Ns_for_table <- lapply(x, N_fun)

means_INT_fun <- function(x){
  mean(x$moral_identity_symbolic)
}
mean_int_for_table <- lapply(x, means_INT_fun)
sd_INT_fun <- function(x){
  sd(x$moral_identity_symbolic)
}
sd_int_for_table <- lapply(x, sd_INT_fun)

means_age_fun <- function(x){
  mean(x$age)
}
mean_age_for_table <- lapply(x, means_age_fun)
sd_age_fun <- function(x){
  sd(x$age)
}
sd_age_for_table <- lapply(x, sd_age_fun)

```

```{r}
test <- 
  cbind.data.frame(
     do.call(rbind,mean_age_for_table)
    ,do.call(rbind,sd_age_for_table)
    ,do.call(rbind,mean_int_for_table)
    ,do.call(rbind,sd_int_for_table)
    ,round(do.call(rbind, rs_for_table),digits=3)
    ,do.call(rbind, ps_for_table)
    ,round(do.call(rbind, Ns_for_table))
    )

test <- `colnames<-`(test, c("mean age","SD age","mean symbolization", "SD symbolization", "r","p","N"))
test

sum(unlist(ps_for_table)<.05)

test1 <- subset(test, p < .05)
test1

total_sig <- length(test1$`mean age`)

neg <- sum(test1$r <0 )
pos <- sum(test1$r>0)

neg_names <- rownames(test1[which(test1$r<0),])
neg_names_first <- c(paste0(neg_names[1:(length(neg_names)-1)], sep = ", "))
neg_names_end <- c("and " , neg_names[length(neg_names)])
noquote(c(neg_names_first, neg_names_end))

pos_names <- rownames(test1[which(test1$r>0),])
pos_names_first <- c(paste0(pos_names[1:(length(pos_names)-1)], sep = ", "))
pos_names_end <- c("and " , pos_names[length(pos_names)])
print(noquote(c(pos_names_first, pos_names_end)))
toString(noquote(paste0(c(pos_names_first, pos_names_end), collapse = "")))
help("toString")

toString(paste0(pos_names_first,collapse = ""))

noquote(paste0(c(pos_names_first, pos_names_end), collapse = ""))
noquote(paste0(c(neg_names_first, neg_names_end), collapse = ""))

```

Similarly, for symbolization, we conducted a linear-mixed-effects model to test if the relationship between age and symbolization holds when participants are nested within country. Our outcome measure was symbolization and our predictor variable was age; we allowed intercepts to vary by country.
Overall, the model significantly predicted symbolization, and provided a better fit for the data than the baseline model,
$\chi$^2^(`r results_anova$Df[2]`) = `r sprintf("%.2f",results_anova$Chisq[2])`, *p* `r paste(p_report(p1))`. 
There was a significant main effect for age,
*F*(`r aov1$NumDF[1]`, `r sprintf("%.2f",round(aov1$DenDF[1],digits=2))`) = 
`r sprintf("%.2f",f3a)`, *p* `r paste(p_report(p3a))`;
with a significant positive association between age and symbolization scores, $b$ = `r sprintf("%.3f",results_coef$Estimate[2])`,
*t*(`r sprintf("%.2f",results_coef$df[2])`) = 
`r sprintf("%.2f",t1)`,
*p* `r paste(p_report(p2))`.
Again, to interpret this, we conducted a series of correlations to test if age was related to symbolization across all countries.  There was a significant correlation between age and symbolization in `r total_sig` countries. Of these, there was a positive correlation between age and symbolization in `r pos` countries, and a negative correlation in `r neg` countries. The countries with a positive correlation were `r noquote(paste0(c(pos_names_first, pos_names_end), collapse = ""))`, and the countries with a negative correlation were `r noquote(paste0(c(neg_names_first, neg_names_end), collapse = ""))` (see Supplementary Materials [Table S_X](https://cillianmiltown.github.io/moral_ID_gender/drafting_supplementary_analysis.html#Age__Symbolization)).